Enhancing \Hinf{} Norm Estimation using Local LPM/LRM Modeling: Applied to an AVIS\thanksref{thx}

### 2016-05-09 ch04-lrmhinf/lrmhinf.tex
Some common commands introduced

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
\begin{equation}
 \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \svdMax{\Delta(\omega)}
  \text{.}
  \label{eq:lrmhinf:omega-star}
\end{equation}
If the measurement time is finite, non-parametric approaches only give access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid $, i.e. $\DFTGrid \isdef \set{\omega_1, \omega_2, \ldots, \omega_{N-1}}$.
Typically, this discrete grid is the frequency grid of the \gls{DFT}.
As such, \eqref{eq:lrmhinf:HinfNorm} can be rewritten as
\begin{equation}
  \gamma \isdef 
    \sup_{\omega} 
      \svdMax{\Delta(\omega)}
         = \max 
             \Set{
                \max_{\omega \in \DFTGrid  } 
                  \svdMax{\Delta(\omega)}
                \;,\;
                \sup_{\omega \in \InterGrid} 
                  \svdMax{\Delta(\omega)}
                }
  \label{eq:lrmhinf:gamma-split-grid}
  \text{.}
\end{equation}
For convenience, we extract both contributions above as:
\begin{align}
  \gammaAtGrid & \isdef 
    \max_{\omega \in \DFTGrid  } 
      \svdMax{\Delta(\omega)}
  \label{eq:lrmhinf:gamma:atGrid}\\
  \gammaIntergrid & \isdef 
    \sup_{\omega \in \InterGrid} 
      \svdMax{\Delta(\omega)}
   \label{eq:lrmhinf:gamma:interGrid}
\end{align}
where $\Omega \isdef \OpenInterval{0,\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Direct inspection of the \gls{FRF} data only allows access to $\gammaAtGrid$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\svdMax{\Delta(\omega_{\star})} = \gamma =\gamma_{\mathrm{FRF}}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\svdMax{\Delta(\omega_{\star})}  = \gamma = \gammaIntergrid$ is determined by the inter-grid behavior.
\end{enumerate}

The use of $\gamma_{\mathrm{FRF}}$ is standard in \gls{FRF}-based approaches, e.g.~\citet{vandeWal2002}.
However, when the measurement time is finite, the expected value of  $\gamma_{\mathrm{FRF}}$  for different realizations of the input signal and noise, may underestimate the actual peak gain $\gamma$, i.e. $\expectedValue{\gammaAtGrid} \leq \gamma$.
This is due to the limited frequency resolution imposed by the measurement time: the frequency where the actual peak gain occurs might not coincide with the excited frequency grid.
Such an underestimate of $\gamma$ leads to overly optimistic controller designs which in turn can cause unstable behavior of the plant.

@@

\figref{fig:lrmhinf:exampleIntergrid} illustrates for $N=97$ that $\gamma_{\mathrm{FRF}}$ for the \gls{SA}, $\lpm{5,4,4}$ and $\lrm{5,2,2,2}$ approaches yield unreliable estimates of $\infnorm{\Delta}$, since the frequency resolution is too coarse.
The specific orders of  \gls{LRM}/\gls{LPM} have been chosen to  keep $N_W$ small while not restricting ourselves to local linear models ($\order{B}=1$). 
Slightly different orders (e.g. $\lrm{4,1,1,1}$) yield comparable results.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/example-both.tikz}
  \caption[Example: \glsentryshort{FRF} underestimates $\E{\gamma}$]{Simulation example revealing that $\gamma_{\mathrm{FRF}}$ (left) for both the \glsfirst{SA}~\legref{leg:example-SA} and \gls{LRM}~\legref{leg:example-LRM} underestimate the true $\gamma = \infnorm{\Delta} = 24$~\legref{leg:example-true-Hinf} due to limited frequency resolution.
  Using the proposed \gls{LRM}-based approach, both the at-grid~($\gamma_{\mathrm{FRF}}$) and intergrid~($\gammaIntergrid$) estimate are improved.
  The interpolated \gls{LRM}~\legref{leg:example-LRM-interpol} and true $\Delta$~\legref{leg:example-true} almost coincide.
  Hence, $\widehat{\infnorm{\Delta_{\LRM}}}_{\infty}$~\legref{leg:example-LRM-interpol-Hinf} approaches the theoretical $\gamma=24$.}
  \label{fig:lrmhinf:exampleIntergrid}
\end{figure}

@@

We simulate the system \eqref{eq:lrmhinf:O2sysDT} for $16\,100$ samples and discard the first $2\tau=100$ samples to randomize the initial conditions.
To mimic the effect of progressively measuring longer data records, the first remaining $N$ samples are used to estimate $\infnorm{\Delta}$ as in \secref{sec:lrmhinf:LPMHinf}.
The values of $N \in \ClosedInterval{35,16\,000}$ are chosen such that $\expectedValue{\gamma_{\mathrm{FRF}}(N)}$ is a non-decreasing function of $N$.
This ensures that grid misalignment between the actual resonance frequency and the observed frequency grid is accounted for such that the worst-case estimate of the resonance is obtained.
These simulations are then repeated in a Monte Carlo simulation of $n_{\mathrm{MC}}=200$ runs where each run has a different realization of  $u_{\Delta}(n)$ and $v(n)$.
The resulting sample average and standard deviation $\sigma$ are shown in \figref{fig:lrmhinf:MC-run-length}.

@@

\paragraph*{Peak value estimates}
\label{sec:lrmhinf:avis-peak-value}
Near $296\unit{Hz}$, a modest improvement of the peak estimate from $-15.3 \unit{dB}$ (at-grid) to $-14.0 \unit{dB}$ is obtained by the interpolation.
The estimate near $\omega_{\star} \approx 328\unit{Hz}$ is improved more substantially.
The proposed method \eqref{eq:lrmhinf:gammaInterpolOptim} yields $\gammaIntergrid=-2.5\unit{dB}$.
$\gamma_{\mathrm{FRF}}\approx -9.9\unit{dB}$ for the estimation data, which means that a $7.4\unit{dB}$ improvement is achieved by the proposed method.
This peak agrees well with the validation data, where the nearest (at-grid) value has an amplitude of $-3.2\unit{dB}$ (i.e. $8\%$ difference).

>>>>

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
\begin{equation}
 \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \svdMax{\Delta(\omega)}
  \text{.}
  \label{eq:lrmhinf:omega-star}
\end{equation}
If the measurement time is finite, non-parametric approaches only give access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid $, i.e. $\DFTGrid \isdef \set{\omega_1, \omega_2, \ldots, \omega_{N-1}}$.
Typically, this discrete grid is the frequency grid of the \gls{DFT}.
As such, \eqref{eq:lrmhinf:HinfNorm} can be rewritten as
\begin{equation}
  \gamma \isdef 
    \sup_{\omega} 
      \svdMax{\Delta(\omega)}
         = \max 
             \Set{
                \max_{\omega \in \DFTGrid  } 
                  \svdMax{\Delta(\omega)}
                \;,\;
                \sup_{\omega \in \InterGrid} 
                  \svdMax{\Delta(\omega)}
                }
  \label{eq:lrmhinf:gamma-split-grid}
  \text{.}
\end{equation}
For convenience, we extract both contributions above as:
\begin{align}
  \gammaAtGrid & \isdef 
    \max_{\omega \in \DFTGrid  } 
      \svdMax{\Delta(\omega)}
  \label{eq:lrmhinf:gamma:atGrid}\\
  \gammaIntergrid & \isdef 
    \sup_{\omega \in \InterGrid} 
      \svdMax{\Delta(\omega)}
   \label{eq:lrmhinf:gamma:interGrid}
\end{align}
where $\Omega \isdef \OpenInterval{0,\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Direct inspection of the \gls{FRF} data only allows access to $\gammaAtGrid$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\svdMax{\Delta(\omega_{\star})} = \gamma =\gamma_{\FRF}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\svdMax{\Delta(\omega_{\star})}  = \gamma = \gammaIntergrid$ is determined by the inter-grid behavior.
\end{enumerate}

The use of $\gamma_{\FRF}$ is standard in \gls{FRF}-based approaches, e.g.~\citet{vandeWal2002}.
However, when the measurement time is finite, the expected value of  $\gamma_{\FRF}$  for different realizations of the input signal and noise, may underestimate the actual peak gain $\gamma$, i.e. $\expectedValue{\gammaAtGrid} \leq \gamma$.
This is due to the limited frequency resolution imposed by the measurement time: the frequency where the actual peak gain occurs might not coincide with the excited frequency grid.
Such an underestimate of $\gamma$ leads to overly optimistic controller designs which in turn can cause unstable behavior of the plant.

@@

\figref{fig:lrmhinf:exampleIntergrid} illustrates for $N=97$ that $\gamma_{\FRF}$ for the \gls{SA}, $\lpm{5,4,4}$ and $\lrm{5,2,2,2}$ approaches yield unreliable estimates of $\infnorm{\Delta}$, since the frequency resolution is too coarse.
The specific orders of  \gls{LRM}/\gls{LPM} have been chosen to  keep $N_W$ small while not restricting ourselves to local linear models ($\order{B}=1$). 
Slightly different orders (e.g. $\lrm{4,1,1,1}$) yield comparable results.

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/example-both.tikz}
  \caption[Example: \glsentryshort{FRF} underestimates $\E{\gamma}$]{Simulation example revealing that $\gamma_{\FRF}$ (left) for both the \glsfirst{SA}~\legref{leg:example-SA} and \gls{LRM}~\legref{leg:example-LRM} underestimate the true $\gamma = \infnorm{\Delta} = 24$~\legref{leg:example-true-Hinf} due to limited frequency resolution.
  Using the proposed \gls{LRM}-based approach, both the at-grid~($\gamma_{\FRF}$) and intergrid~($\gammaIntergrid$) estimate are improved.
  The interpolated \gls{LRM}~\legref{leg:example-LRM-interpol} and true $\Delta$~\legref{leg:example-true} almost coincide.
  Hence, $\widehat{\infnorm{\Delta_{\LRM}}}_{\infty}$~\legref{leg:example-LRM-interpol-Hinf} approaches the theoretical $\gamma=24$.}
  \label{fig:lrmhinf:exampleIntergrid}
\end{figure}

@@

We simulate the system \eqref{eq:lrmhinf:O2sysDT} for $16\,100$ samples and discard the first $2\tau=100$ samples to randomize the initial conditions.
To mimic the effect of progressively measuring longer data records, the first remaining $N$ samples are used to estimate $\infnorm{\Delta}$ as in \secref{sec:lrmhinf:LPMHinf}.
The values of $N \in \ClosedInterval{35,16\,000}$ are chosen such that $\expectedValue{\gamma_{\FRF}(N)}$ is a non-decreasing function of $N$.
This ensures that grid misalignment between the actual resonance frequency and the observed frequency grid is accounted for such that the worst-case estimate of the resonance is obtained.
These simulations are then repeated in a Monte Carlo simulation of $n_{\mathrm{MC}}=200$ runs where each run has a different realization of  $u_{\Delta}(n)$ and $v(n)$.
The resulting sample average and standard deviation $\sigma$ are shown in \figref{fig:lrmhinf:MC-run-length}.

@@

\paragraph*{Peak value estimates}
\label{sec:lrmhinf:avis-peak-value}
Near $296\unit{Hz}$, a modest improvement of the peak estimate from $-15.3 \unit{dB}$ (at-grid) to $-14.0 \unit{dB}$ is obtained by the interpolation.
The estimate near $\omega_{\star} \approx 328\unit{Hz}$ is improved more substantially.
The proposed method \eqref{eq:lrmhinf:gammaInterpolOptim} yields $\gammaIntergrid=-2.5\unit{dB}$.
$\gamma_{\FRF}\approx -9.9\unit{dB}$ for the estimation data, which means that a $7.4\unit{dB}$ improvement is achieved by the proposed method.
This peak agrees well with the validation data, where the nearest (at-grid) value has an amplitude of $-3.2\unit{dB}$ (i.e. $8\%$ difference).
